---
title: "textstem"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  md_document:
    toc: true      
---

```{r, echo=FALSE}
desc <- suppressWarnings(readLines("DESCRIPTION"))
regex <- "(^Version:\\s+)(\\d+\\.\\d+\\.\\d+)"
loc <- grep(regex, desc)
ver <- gsub(regex, "\\2", desc[loc])
verbadge <- sprintf('<a href="https://img.shields.io/badge/Version-%s-orange.svg"><img src="https://img.shields.io/badge/Version-%s-orange.svg" alt="Version"/></a></p>', ver, ver)

pacman::p_load_current_gh('trinker/numform')
nr <- numform::f_comma(length(presidential_debates_2012$dialogue))
nw <- numform::f_comma(sum(stringi::stri_count_words(presidential_debates_2012$dialogue), na.rm = TRUE))
````

[![Build Status](https://travis-ci.org/trinker/textstem.svg?branch=master)](https://travis-ci.org/trinker/textstem)
[![Coverage Status](https://coveralls.io/repos/trinker/textstem/badge.svg?branch=master)](https://coveralls.io/r/trinker/textstem?branch=master)
`r verbadge`

**textstem** is a tool-set for stemming and lemmatizing words.  Stemming is a process that removes affixes.  Lemmatization is the process of reducing words to base form.  

# Functions

The main functions, task category, & descriptions are summarized in the table below:


| Function                      | Task        | Description                                | 
|-------------------------------|-------------|--------------------------------------------| 
| `stem_words`                  | stemming    | Stem words                                 | 
| `stem_strings`                | stemming    | Stem strings                               | 
| `lemmatize_words`             | lemmatizing | Lemmatize words                            | 
| `lemmatize_strings`           | lemmatizing | Lemmatize strings                          | 
| `make_lemma_dictionary_words` | lemmatizing | Generate a dictionary of lemmas for a text |     


# Installation

To download the development version of **textstem**:

Download the [zip ball](https://github.com/trinker/textstem/zipball/master) or [tar ball](https://github.com/trinker/textstem/tarball/master), decompress and run `R CMD INSTALL` on it, or use the **pacman** package to install the development version:

```r
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("trinker/textstem")
```

# Contact

You are welcome to:    
- submit suggestions and bug-reports at: <https://github.com/trinker/textstem/issues>    
- send a pull request on: <https://github.com/trinker/textstem/>    
- compose a friendly e-mail to: <tyler.rinker@gmail.com>    


# Examples

The following examples demonstrate some of the functionality of **textstem**.

## Load the Tools/Data

```{r, message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(textstem, dplyr)

data(presidential_debates_2012)
```


## Stemming

Stemming is the act of removing inflections from a word not necessarily ["identical to the morphological root of the word" (wikipedia).](https://en.wikipedia.org/wiki/Stemming).  Below I show stemming of several small strings.

```{r}
y <- c(
    'the dirtier dog has eaten the pies',
    'that shameful pooch is tricky and sneaky',
    "He opened and then reopened the food bag",
    'There are skies of blue and red roses too!',
    NA,
    "The doggies, well they aren't joyfully running.",
     "The daddies are coming over...",
    "This is 34.546 above"
)
stem_strings(y)
```

## Lemmatizing

Lemmatizing is the ["grouping together the inflected forms of a word so they can be analysed as a single item" (wikipedia)](https://en.wikipedia.org/wiki/Lemmatisation).  In the example below I reduce the strings to their lemma form.  `lemmatize_strings` uses a lookup dictionary.  The default uses [Mechura's (2016)](http://www.lexiconista.com) English lemmatization list. The `make_lemma_dictionary` function contains two additional engines for generating a lemma lookup table for use in `lemmatize_strings`.


```{r}
y <- c(
    'the dirtier dog has eaten the pies',
    'that shameful pooch is tricky and sneaky',
    "He opened and then reopened the food bag",
    'There are skies of blue and red roses too!',
    NA,
    "The doggies, well they aren't joyfully running.",
     "The daddies are coming over...",
    "This is 34.546 above"
)
lemmatize_strings(y)
```

This lemmatization uses the [**hunspell**](https://CRAN.R-project.org/package=hunspell) package to generate lemmas.


```{r}
lemma_dictionary_hs <- make_lemma_dictionary(y, engine = 'hunspell')
lemmatize_strings(y, dictionary = lemma_dictionary_hs)
```

This lemmatization uses the [**koRpus**](https://CRAN.R-project.org/package=koRpus) package and the [TreeTagger](http://www.cis.uni-muenchen.de/~schmid/tools/TreeTagger/) program to generate lemmas.

```{r}
lemma_dictionary_tt <- make_lemma_dictionary(y, engine = 'treetagger')
lemmatize_strings(y, lemma_dictionary_tt)
```


It's pretty fast too.  Observe:

```{r}
tic <- Sys.time()

presidential_debates_2012$dialogue %>%
    lemmatize_strings() %>%
    head()

(toc <- Sys.time() - tic)
```


That's `r nr` rows of text, or `r nw` in `r round(as.numeric(toc), 2)` seconds.






